Fractional Types

### 2019-11-21 fracGC/p.tex
example execution traces

\input{latex/PiFracDynDefCode.tex}
\input{latex/PiFracDynCode.tex}
\input{latex/BooleanCircuitsCode.tex}
\input{latex/PiPointedFracCode.tex}
\input{latex/ExtractionCode.tex}
\input{latex/PiMemSemCode.tex}
\input{latex/PiFracMemSemCode.tex}
\input{latex/J.tex}

@@

There are two main reasons we are calling the new type constructor a
\emph{fractional type} and using the suggestive notation
$\oneover{t}$:
\begin{itemize}
\item Categorical: ... explain and point to section with examples
  showing 1/ab = 1/a  1/b etc.
\item Operational ... this section; bits = log of size; fractional
  needs negative bits; i.e. space collector; make this precise  
\end{itemize}

@@

>>>>

%% for Sec. 4
\input{latex/PiFracDynDefCode.tex}
\input{latex/PiFracDynCode.tex}

%% for Sec. 5
%% \input{latex/PiMemSemCode.tex}
\input{latex/PiFracMemSemCode.tex}

%% for Sec. 6
\input{latex/J.tex}
\input{latex/PiPointedFracCode.tex}
\input{latex/ExtractionCode.tex}
\input{latex/BooleanCircuitsCode.tex}

@@

There are two technical reasons we are calling the new type constructor a
\emph{fractional type} and hence using for it the suggestive notation
$\oneover{t}$:
\begin{itemize}
\item Categorical: The categorical model for the plain $\Pi$ language
  of type isomorphisms is a category with two symmetric monoidal
  structures (additive and multiplicative) with the multiplicative
  fragment distributing over the additive one~\cite{XXX}. Generally,
  a symmetric monoidal category is \emph{compact closed} if every
  object has a dual object whose behavior is governed by our $\eta$
  and $\epsilon$ operators~\cite{XXX}. Our language from the previous
  section is an adaptation of this idea to the context of $\Pi$. Many
  programming abstractions (such as \emph{name}, \emph{co-name}, and
  \emph{trace} illustrated in Sec.~\ref{sec:cat}) can be automatically
  derived from properties of compact closed categories. Additionally,
  all the properties expected of the rational numbers
  (e.g. $\oneover{(m*n)} = \oneover{m}*\oneover{n}$,
  $\oneover{(\oneover{n})}=n$, etc.) can be given constructive content
  corresponding to manipulation of gc processes. We have already seen
  the first example and will see more in Sec.~\ref{sec:cat}.

\item Operational: The other reason our types can be justifiably
  called fractional is information theoretic. As mentioned in the
  introduction the number of bits needed to store information about a
  space with $n$ distinct values is $\log{n}$. If the number of bits
  can be thought of as a debt (to be reclaimed) then a negative number
  of bits corresponds to a space with $\oneover{n}$ distinct
  values. Indeed we interpret these values as gc processes each
  specialized to collect space for one of the $n$ distinct values. In
  this section, we make this intuition precise by presenting a
  (fully-formalized in Agda) abstract machine that manages an explicit
  heap. 
\end{itemize}

@@

CNOT trace in plain pi:

\includegraphics[scale=0.4]{trace1.png}

\PIFMEMsize{}

If a type has size $n$ we will use a number $0..(n-1)$ to represent
the values of the types. In binary, this would use $\log{n}$ bits.

The abstract machine has a register representing the code and a
pointer. If the code is acting on a type with $n$ values, the pointer
has $\log{n}$ bits (rounded).

In plain pi, if the input is representable using $n$ bits then this is
the only memory you get to use throughout the entire circuit. The
circuit just transforms these $n$ bits; example.

In frac pi, you are however allowed to use more than $n$ bits in the
intermediate steps as eta is used to allocate more space.

Use Fin size as proxy for memory size

zigzag trace

\includegraphics[scale=0.5]{trace2.png}
